---
title: "Replication of Study 3 by Stavrova, Ehlebracht & Vohs (2020, Journal of Experimental Psychology)"
author: "Eric Neumann (ericnm@stanford.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
---


##Introduction


### Background and Replication Choice

Cynicism is highly prevalent in the US these days, and most of its correlates are negative. One relevant question is thus why it manages to sustain itself in the light of its many disadvantages. Stavrova et al (2020) propose an interesting self-reinforcing circle of cynicism: being cynical makes you treat others with more disrespect, and as they will treat you with more disrespect in turn, you are more likely to be cynical. Study 3 of their paper examines this latter part, finding that recalling an experience of being disrespected leads people to be more cynical about human nature. As much of my PhD research will likely build on this finding, I would like to first see if the results are robust and replicable.

### Stimuli and Procedures

Participants were recruited on Prolific.com. They were first asked to write a short essay about a time they had experienced disrespect (experimental condition) or guilt (control condition). This was followed by a manipulation check where participants indicated the extent to which they felt disrespected or guilty. All subjects then completed the cynical distrust scale (Greenglass & Julkunen, 1989). As participants might have seen through the experimental design, they also filled out the Perceived Awareness of the Research Hypothesis scale (PARH; Rubin, 2010). Finally, they completed an attention check and basic demographic questions. 

### Anticipated Challenges

I anticipate my primary challenge to be the coding of the analysis of the PARH scale as the original paper proposed four different ways perceived awareness can be measured. I do think, however, that learning this will be useful for future studies and also estimate the rest of the analysis to be straightforward enough to allow for time exploring the impact of participant's perceived awareness of the intentions behind the study.

### Pertinent Links

This replication attempt is documented in Github [here.](https://github.com/psych251/stavrova2020)
The original paper, within the same repository, can be found [here.](https://github.com/psych251/stavrova2020/blob/master/original_paper/paper.pdf)


##Methods


###Power Analysis

Original effect size, power analysis for samples to achieve 80%, 90%, 95% power to detect that effect size. Considerations of feasibility for selecting planned sample size.

###Planned Sample

MTurk participants in the United States will be allowed to partake in this study. There are no specific preselection criteria.

###Materials

Prompts about experience making ss feel guilty or disrespected
Cynical Distrust Scale (Greenglass & Julkunen, 1989)
Perceived Awareness of the Research Hypothesis scale (PARH; Rubin, 2010)
Demographics

###Procedure	

Participants are first prompted to write about an experience where they felt guilt or disrespect.

[guilt] "In this study, we will ask you to describe events in your life. Please describe in detail a time in which you felt guilty over something that happened between you and another person. Indicate where and when this happened and provide the initials of that person."

[disrespect] "In this study, we will ask you to describe events in your life. Please describe in detail a time when you felt disrespected by someone else. Perhaps you felt like you were treated with less courtesy than others, or you were slighted by someone, for example. Please, indicate where and when this happened and provide the initials of the person(s) who treated you that way, if possible."

As a manipulation check, all participants then answer if recalling this experience makes them feel guilty or disrespected.

"Does reliving the experience you wrote about make you feel guilty right now?" 1 (not at all) - 7 (very much)
"Does reliving the experience you wrote about make you feel disrespected right now?" 1 (not at all) - 7 (very much)

Then, on a scale from 1 (not at all) to 7 (very much), participants respond to the cynical distrust scale.

"Please indicate to what extent you agree or disagree with the following statements.
1. I think most people would lie to get ahead.
2. Most people inwardly dislike putting themselves out to help other people.
3. Most people make friends because friends are likely to be useful to them.
4. It is safer to trust nobody.
5. No one cares much what happens to you.
6. Most people are honest chiefly through fear of being caught.
7. I commonly wonder what hidden reasons another person may have for doing something
nice to me.
8. Most people will use somewhat unfair means to gain profit or an advantage rather than lose it."
9 When a man is with a woman he is usually thinking about things related to her sex

Unlike in the original study, the ninth item from the original survey is not omitted for face validity reasons.

Like in the original study, an attention check is included.

"To monitor data quality, please select the middle option here."

Next, participants fill out the PARH scale.

"1. I knew what the researchers were investigating in this research. 
2. I wasn’t sure what the researchers were trying to demonstrate in this research. 
3. I had a good idea about what the hypotheses were in this research. 
4. I was unclear about exactly what the researchers were aiming to prove in this research."

Finally, participants are asked certain demographic questions.

"Please, indicate your gender. (male - female) [I will also include a third category 'other, prefer not to say']
How old are you?"

I will include a few additional demographic variables for exploratory purposes.

"When it comes to politics, would you describe yourself as liberal, conservative, or neither liberal nor conservative?" (adapted from Guess et al, 2020) (scale from 'very liberal' to 'very conservative')
"What is the highest level of education you have completed?" (scale from 'did not graduate from high school' to 'postgraduate')

I run the replication on mTurk and will thus omit the question about the Prolific ID. The study concludes with

"Thank you for your participation."

###Analysis Plan

Manipulation check - independent t-tests: do participants in guilt condition feel more guilt? do participants in disrespect condition feel more disrespect?

Cynicism (key measure) - independent t-test: are participants in disrespect condition more cynical than those in guilt condition?

Demand effects of PARH scale - four different tests: compare participant PARH scale responses to theoretical midpoint (one sample t-test); exclude positive outliers (+3SDs), test associations between PARH scores and IVs and DVs (r(PARH, cynicism), interaction condition x PARH, independent t-test between conditions on PARH score), include PARH score as covariate in main hypothesis (ANCOVA).

###Differences from Original Study

This replication will be run on mTurk, not Prolific. I will include the ninth item of the Cynical Distrust scale despite its low face validity. The gender demographics question includes a third option, and two demographic questions (political leaning, education) have been added strictly for exploratory purposes. I do not expect any of these changes to impact the results.

### Methods Addendum (Post Data Collection)

You can comment this section out prior to final report with data collection.

#### Actual Sample
  Sample size, demographics, data exclusions based on rules spelled out in analysis plan

#### Differences from pre-data collection methods plan
  Any differences from what was described as the original plan, or “none”.


##Results


### Data preparation

Data preparation following the analysis plan.

```{r include=T}
###load libraries, likely only need tidyverse packages
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

```{r include=T}
###import data
testdata <- read.csv("testdata.csv")
View(testdata)
```

```{r include=T}

###delete rows where I test-ran study
testdata_analysis <- testdata[-c(1:8), ]
View(testdata_analysis)
```

```{r include=T}
###exclude all who failed attention check
testdata_analysis_excl <- testdata_analysis %>%
  filter(att_check == "4")
View(testdata_analysis_excl)
```

```{r include=T}
###create aggregate cynicism score
testdata_analysis_excl[, 22:30] <- sapply(testdata_analysis_excl[, 22:30], FUN = as.numeric)
testdata_analysis_excl$cyn_total <- (testdata_analysis_excl$cyn1 + testdata_analysis_excl$cyn2 + testdata_analysis_excl$cyn3 + 
                                       testdata_analysis_excl$cyn4 + testdata_analysis_excl$cyn5 + testdata_analysis_excl$cyn6 + 
                                       testdata_analysis_excl$cyn7 + testdata_analysis_excl$cyn8 + testdata_analysis_excl$cyn9)/9
summary(testdata_analysis_excl$cyn_total)
```

```{r include=T}
###separate by condition
cyn_means <- testdata_analysis_excl %>%
  group_by(cond_disr) %>% #for actual analysis, first create column for value Y/N in cond_guilt column, then group by conditions
  summarise (cyn_total = mean(cyn_total))
View(cyn_means)
```

### Confirmatory analysis

```{r include=T}
###run t-test on cynicism score between two conditions
cyn_means_guilt <- cyn_means[1, ]
cyn_means_disr <- cyn_means[2, ]
#t.test(cyn_means_guilt$cyn_total, cyn_means_disr$cyn_total, paired = F) --> error because I only have 2 data points :)
```

###Exploratory analyses

Any follow-up analyses desired (not required).  

## Discussion

### Summary of Replication Attempt

Open the discussion section with a paragraph summarizing the primary result from the confirmatory analysis and the assessment of whether it replicated, partially replicated, or failed to replicate the original result.  

### Commentary

Add open-ended commentary (if any) reflecting (a) insights from follow-up exploratory analysis, (b) assessment of the meaning of the replication (or not) - e.g., for a failure to replicate, are the differences between original and present study ones that definitely, plausibly, or are unlikely to have been moderators of the result, and (c) discussion of any objections or challenges raised by the current and original authors about the replication attempt.  None of these need to be long.
